Stack Safety is a Security Property

### 2021-04-10 tex/main.tex
recover most of old intro as comments

So how should we formally define stack safety? Definitions in the literature are commonly informal
and proceed by enumerating the specific bad things that a given enforcement mechanism
prevents, such as corruption of return addresses, buffer overflows, or
use of uninitialized variables.
%\bcp{Does use of unitialized variables lead to stack safety attacks?\apt{sure, via buffer overflow}}
%with the notion usually being
%defined negatively: through examples of stack-based exploits
%\rb{usually though not always; as noted below there are other
%  approaches}\bcp{Yes, we have to be very careful about this}.
%\leo{Examples at a high level}: such as -- does not overwrite return addresses, or ...
The most abstract characterization of which we are aware is in 
recent work by \citet{Skorstengaard+19}, which defines stack safety as the conjunction of two properties:
{\em local state encapsulation (LSE)} and {\em well-bracketed control flow (WBCF)}.
Informally, LSE says that the contents of
a caller's stack frame are not read and written while its callees (and their callees, etc.) are
executing, and WBCF says that callees always
return to the point just after their call (or don't return at all). 
\citet{Skorstengaard+19} formalize these properties by defining an idealized machine
in which they visibly hold by construction; an enforcement mechanism (in their case, hardware capabilities)
can then be shown to provide stack safety by giving a fully abstract embedding function
from the idealized machine to an ordinary machine equipped with the enforcement.

@@

While our property definitions were developed with micro-policy enforcement in mind, they
agnostic about many implementation details and should be applicable to other
systems with minor adjustment.

%\begin{figure}[b]
%    \begin{minipage}{.4\textwidth}
%\begin{verbatim}
%int main() {
%  int x = 42;
%  f();
%  print(x);
%}
%\end{verbatim}
%    \end{minipage}
%    \begin{minipage}{.4\textwidth}
%\begin{verbatim}
%int main() {
%  int x = 1;
%  f();
%  if(x)
%    g();
%  else
%    h();
%}
%\end{verbatim}
%    \end{minipage}
%  \caption{Simple Examples}
%  \label{fig:examples1}
%\end{figure}

%\begin{figure}[b]
%    \begin{minipage}{.4\textwidth}
%\begin{verbatim}
%int main(int arg) {
%  if(arg)
%    g();
%  else
%    h();
%  f();
%}
%\end{verbatim}
%    \end{minipage}
%    \begin{minipage}{.4\textwidth}
%\begin{verbatim}
%int main() {
%  int x = 1;
%  f();
%  x = -x;
%  f();
%  print(x);
%}
%\end{verbatim}
%  \end{minipage}
%
%  \caption{More Examples}
%  \label{fig:examples2}
%\end{figure}

In summary, we offer the following contributions:
\begin{itemize}
\item 
  We harness the language-based security concepts of integrity and confidentiality 
  to formalize a parameterized {\em stepwise stack-safety property}, and argue that
  it captures the intuitive meaning of local stack encapsulation, well-bracketed control
  flow, and temporal stack safety~(\cref{sec:lse-and-wbcf}).
\item
  We apply a property-based randomized testing framework to provide evidence that 
  this stepwise property is enforced by the Depth Isolation policy of
  \citet{DBLP:conf/sp/RoesslerD18} (\cref{sec:conservative}).
\item 
  We formalize a weaker but more extensional \emph{observational} variant of the
  stack safety property, and argue that it accurately characterizes 
  the Lazy Tagging and Clearing policy of \citet{DBLP:conf/sp/RoesslerD18},
  when enhanced with unique tags on each activation~(\cref{sec:lazy}).
\item 
  We show how our models can be instantiated and extended to handle passing
  arguments on the stack~(\cref{sec:??}) and to support simple coroutining~(\cref{sec:??}).
%  This policy does not enforce our strict
%  local state encapsulation property,
%  but a modified version of it using unique per-activation tags enforces
%  an (also provably) weaker lazy variant of the property, which still prohibits
%  the same intuitively stack unsafe behaviors.
%
% \apt{Awkward, but the truth I fear.} \bcp{I agree the phrasing is awkward,
%   but the point we can make is still strong: To make the property easier
%   to enforce, we can slightly weaken its guarantees; the weaker version is
%   a bit more complicated to state but still retains the pragmatic
%%   protections of the original.}\apt{To clarify: Nick's LT\&C policy uses
%   depth-based tags, which do not enforce even our lazy policy. So we need to
%   introduce a version of LT\&C that uses unique activation tags.}

>>>>

So how should we formally define stack safety? Definitions in the literature are commonly informal
and proceed by enumerating the specific bad things that a given enforcement mechanism
prevents, such as corruption of return addresses, buffer overflows, or
use of uninitialized variables.
%\bcp{Does use of unitialized variables lead to stack safety attacks?\apt{sure, via buffer overflow}}
%with the notion usually being
%defined negatively: through examples of stack-based exploits
%\rb{usually though not always; as noted below there are other
%  approaches}\bcp{Yes, we have to be very careful about this}.
%\leo{Examples at a high level}: such as -- does not overwrite return addresses, or ...
The most abstract characterization of which we are aware is in 
recent work by \citet{Skorstengaard+19}, which defines stack safety as the conjunction of two properties:
{\em local state encapsulation (LSE)} and {\em well-bracketed control flow (WBCF)}.
Informally, LSE says that the contents of
a caller's stack frame are not read and written while its callees (and their callees, etc.) are
executing, and WBCF says that callees always
return to the point just after their call (or don't return at all). 
\citet{Skorstengaard+19} formalize these properties by defining an idealized machine
in which they visibly hold by construction; an enforcement mechanism (in their case, hardware capabilities)
can then be shown to provide stack safety by giving a fully abstract embedding function
from the idealized machine to an ordinary machine equipped with the enforcement.
%%
%% While this work represents an important step in the right direction, we believe
%% there is still room to craft an operational characterization
%% of stack safety. \apt{Perhaps our rhetoric should be more micro-policy-specific here?}
%% In other words, we seek formal criteria that can be applied
%% to a mechanism for protecting stack-based low-level programs to judge
%% whether or not it actually ``guarantees stack safety.''
%% \rb{The remainder of this paragraph may be diving too quickly, would a reader be
%% confused by the references to these types of properties, yet to be introduced?}
%% We begin with a strong
%% local state encapsulation property that closely models the behavior of eager
%% micro-policies from \citet{DBLP:conf/sp/RoesslerD18}. This ``eager'' property
%% is too strong to characterize stack safety; reasonable enforcement mechanisms
%% exist that violate it. One such mechanism is a ``lazy'' tag policy, similar to
%% the one in \citet{DBLP:conf/sp/RoesslerD18}. We weaken local state encapsulation
%% to support this technique and argue that the new property, defined relative to an
%% arbitrary notion of observation. This property is still strong enough to prevent
%% the bad behaviors we associate with unsafe stack behavior, but weak enough that
%% other reasonable enforcement mechanisms can enforce it.

@@

While our property definitions were developed with micro-policy enforcement in mind, they
agnostic about many implementation details and should be applicable to other
systems with minor adjustment.\apt{move this up?}

%% %\begin{figure}[b]
%% %    \begin{minipage}{.4\textwidth}
%% %\begin{verbatim}
%% %int main() {
%% %  int x = 42;
%% %  f();
%% %  print(x);
%% %}
%% %\end{verbatim}
%% %    \end{minipage}
%% %    \begin{minipage}{.4\textwidth}
%% %\begin{verbatim}
%% %int main() {
%% %  int x = 1;
%% %  f();
%% %  if(x)
%% %    g();
%% %  else
%% %    h();
%% %}
%% %\end{verbatim}
%% %    \end{minipage}
%% %  \caption{Simple Examples}
%% %  \label{fig:examples1}
%% %\end{figure}

%% %\begin{figure}[b]
%% %    \begin{minipage}{.4\textwidth}
%% %\begin{verbatim}
%% %int main(int arg) {
%% %  if(arg)
%% %    g();
%% %  else
%% %    h();
%% %  f();
%% %}
%% %\end{verbatim}
%% %    \end{minipage}
%% %    \begin{minipage}{.4\textwidth}
%% %\begin{verbatim}
%% %int main() {
%% %  int x = 1;
%% %  f();
%% %  x = -x;
%% %  f();
%% %  print(x);
%% %}
%% %\end{verbatim}
%% %  \end{minipage}
%% %
%% %  \caption{More Examples}
%% %  \label{fig:examples2}
%% %\end{figure}

%% \paragraph*{The Use of Stack Safety}

%% How does the programmer rely on stack safety? The stack makes implicit promises about
%% the behavior of function calls. First, that when a call is finished, the caller can continue
%% from where it left off with its local state unchanged. And second, that the call's
%% behavior is determined only by a predictable interface that does not include the caller's
%% local state.

%% %The most natural example is the first
%% %in figure \ref{fig:examples1}, in which {\tt main} calls stores 42 to a local variable {\tt x},
%% %calls {\tt f}, and then prints {\tt x}. Here the programmer should be able to expect that
%% %{\tt main} will print 42, but in C such an assumption may be wrong, if {\tt f} should
%% %overwrite {\tt main}'s stack data.

%% Both expectations imply predictable control flow, termed {\em well-bracketed
%% control flow} (WBCF) by \citeauthor{Skorstengaard+19}. Calls always jump to valid entry
%% points and return to their call sites. Then the data portion of the first expectation is for
%% {\em integrity}: that a caller's local data is protected from modification by its callee.
%% This is vital for local reasoning about the behavior of the caller. The security implications
%% are significant: without integrity, a caller that branches on a local variable could be
%% manipulated by its callee into unexpected behavior.

%% Alongside integrity, we have {\em confidentiality}, in which a caller's data is protected
%% from leakage by its callee. As a reasoning principle, confidentiality tells us that to
%% predict the behavior of a call we need only consider a subset of the machine state.
%% And in security terms, we can guarantee the secrecy of local data that we haven't passed
%% even when calling a badly behaved function. Together, integrity and confidentiality
%% comprise \citeauthor{Skorstengaard+19}'s {\em local state encapsulation} (LSE) property.

%% It must be reiterated that LSE is of little use without WBCF: without WBCF, a callee
%% could subvert LSE by ``returning'' to arbitrary code; after the return LSE should permit
%% accessing the caller's state, but the execution would remain under the callee's control.

%% To phrase LSE and WBCF as security properties requires an explicit attacker model:
%% a clear understanding of who is being protected and from what. Most existing
%% stack protection mechanisms envisage scenarios where a callee is
%% influenced by external inputs to corrupt the stack in some way. To
%% avoid getting into the details of exactly how this happens when
%% formalizing our properties, we use an even stronger model: when we
%% state the protections that a function can rely on when it makes a
%% call, we assume that the callee can execute arbitrary instructions
%% until it returns. Thus we have no need to differentiate safe code
%% from adversarial --- all functions, and indeed all activations of the same function,
%% are mutually distrusting.
%% %
%% For this model to make sense, we need to know when a call or a return is taking
%% place. We assume that calls and returns are annotated in the code, so that
%% we can track the point at which control transfers from the protected caller
%% to the attacking callee --- and when that callee in turn becomes a protected
%% caller in its own right.

%% This ``nesting'' of properties is a key challenge in adapting the concepts
%% of language-based security to the stack setting. We address this issue by
%% turning any set of annotations into a dynamic model of {\em domains},
%% in which some stack locations are sealed by the caller that needs to preserve
%% them while the rest are unsealed and available for use by the active, attacking
%% callee.

%% First we introduce this model for a single simple stack with no sharing between callers and
%% callees. Then we expand it to allow passing of stack data, granting a callee one-time
%% access to the caller's data. We do not model more sophisticated sharing, because to do
%% so realistically would require a separate protection scheme, such as a capability
%% model. While this is quite feasible in a tag-based enforcement mechanism, it is properly
%% a separate policy outside the scope of stack safety. (A tag-based approach can easily
%% compose multiple such policies.)

%% Finally, we extend the model to a coroutine system featuring finitely-bounded stacks.
%% Though simplistic, this extension is a step toward concurrency and demonstrates that
%% the model is flexible enough to handle more sophisticated styles of control-flow.

%% After formally defining stack safety, we show that the definition is
%% enforceable using an existing tag-based enforcement
%% mechanism~\citep{DBLP:conf/sp/RoesslerD18}. We first consider
%% \citeauthor{DBLP:conf/sp/RoesslerD18}'s conservative Depth Isolation policy.
%% This policy testably enforces our {\em eager} properties, in which integrity is respected
%% at every step of a program, and each call respects integrity at each of its internal steps.
%% These properties are very strong, and our policies enforce them; this also makes them very
%% intensional, in that they do not distinguish internal behavior from external.
%% They also have the advantage that they are suitable for efficient testing, as they
%% fail immediately when a violation occurs.

%% The downside of the Depth Isolation policy is that it is rather expensive to implement.
%% In the same work, \citeauthor{DBLP:conf/sp/RoesslerD18} propose a different mechanism
%% that is more efficient by virtue of being lazy in its enforcement.
%% Instead of enforcing the integrity and confidentiality of the stack at the
%% exact moment when a callee returns, this policy signals violations when the
%% caller actually accesses data that the callee has written into its stack
%% frame. \ifaftersubmission\bcp{We could also hint at the lazy confidentiality
%%   property in the same way...}\fi
%% The lazy policy does not enforce our eager integrity property, and to our
%% knowledge there is no formal statement of the properties that it ought to enforce.
%% We propose {\em lazy} stack-safety properties, which capture the desired behavior of a
%% policy that detects integrity violations between when they occur and when they make
%% a visible impact. Intuitively this can be phrased as ``no visible reads after dangerous
%% writes.'' Unfortunately \citeauthor{DBLP:conf/sp/RoesslerD18}'s existing
%% ``Lazy Tagging and Clearing'' policy does not quite enforce this property, as we show in
%% testing, but formalizing the ideal property allows us to see how it falls short.
%% We provide a variant of ``Lazy Tagging and Clearing'' that does enforce lazy integrity
%% at some performance cost.
%% %We prove (in Coq) that the eager property implies the lazy one.\apt{do we?
%% %  and is it so interesting anyhow?}
%% We argue that any system purporting to enforce stack safety ought to obey at least lazy
%% stack safety, making it a plausible extensional characterization.

%% %\bcp{I worry that we are being a little
%% %  dishonest here: the stated advantage of the lazy policy is that it is more
%% %efficient, but our variant is {\em not} efficient!}

In summary, we offer the following contributions:
\begin{itemize}
\item 
  We harness the language-based security concepts of integrity and confidentiality 
  to formalize a parameterized {\em stepwise stack-safety property}, and argue that
  it captures the intuitive meaning of local stack encapsulation, well-bracketed control
  flow, and temporal stack safety~(\cref{sec:lse-and-wbcf}).
\item
  We apply a property-based randomized testing framework to provide evidence that 
  this stepwise property is enforced by the Depth Isolation policy
  of~\citet{DBLP:conf/sp/RoesslerD18}~(\cref{sec:conservative}).
\item 
  We formalize a weaker but more extensional \emph{observational} variant of the
  stack safety property, and argue that it accurately characterizes 
  the Lazy Tagging and Clearing policy of~\citet{DBLP:conf/sp/RoesslerD18},
  when enhanced with unique tags on each activation~(\cref{sec:lazy}).
\item 
  We show how our models can be instantiated and extended to handle passing
  arguments on the stack~(\cref{sec:??}) and to support simple coroutining~(\cref{sec:??}).
%  This policy does not enforce our strict
%  local state encapsulation property,
%  but a modified version of it using unique per-activation tags enforces
%  an (also provably) weaker lazy variant of the property, which still prohibits
%  the same intuitively stack unsafe behaviors.
%
% \apt{Awkward, but the truth I fear.} \bcp{I agree the phrasing is awkward,
%   but the point we can make is still strong: To make the property easier
%   to enforce, we can slightly weaken its guarantees; the weaker version is
%   a bit more complicated to state but still retains the pragmatic
%%   protections of the original.}\apt{To clarify: Nick's LT\&C policy uses
%   depth-based tags, which do not enforce even our lazy policy. So we need to
%   introduce a version of LT\&C that uses unique activation tags.}
